---
title: "Business Interruption EDA"
output: html_document
date: "2026-03-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(dplyr)
library(plotly)
library(tidyr)
library(knitr)
library(writexl)
```


## Missing Values Summary
```{r echo=FALSE}

# Data cleaning steps
claims_data <- read_excel("BI Claims.xlsx")
original_row_count <- nrow(claims_data)

# Display table that shows the number of missing rows
missing_summary <- claims_data %>%
  summarise(across(everything(), ~mean(is.na(.)) * 100)) %>%  # % missing
  pivot_longer(
    cols = everything(),
    names_to = "column",
    values_to = "percent_missing"
  )

# overall number of missing rows
rows_with_na <- claims_data %>%
  filter(if_any(everything(), is.na)) %>%  # TRUE if any column in the row is NA
  nrow()

overall_percent <- rows_with_na / nrow(claims_data) * 100

# add overall row to the summary
missing_summary <- missing_summary %>%
  bind_rows(tibble(
    column = "Rows with NA",
    percent_missing = overall_percent
  ))

# Display table in R Markdown
kable(missing_summary, digits = 2, caption = "Percent of Missing Values per Column")

# cleaning solar system
claims_data <- claims_data %>%
  drop_na() %>%
  mutate(
    policy_id = sub("_.*", "", policy_id),
    station_id = sub("_.*", "", station_id),
    solar_system = sub("_.*", "", solar_system),
    avg_crew_exp = round(avg_crew_exp),
    claim_count = round(claim_count)
  ) %>%
  filter(
    production_load >= 0,
    production_load <= 1,
    energy_backup_score %in% c(1,2,3,4,5),
    safety_compliance %in% c(1,2,3,4,5),
    supply_chain_index >= 0,
    supply_chain_index <= 1,
    avg_crew_exp >= 1,
    avg_crew_exp <= 30,
    maintenance_freq %in% c(0,1,2,3,4,5,6),
    exposure >= 0,
    exposure <= 1,
    claim_count %in% c(0,1,2,3,4)
  )


# Removed Rows
paste0("Removed ", ((original_row_count - nrow(claims_data)) / original_row_count) * 100, "% of the data (defective)")

```
## Distribution of Claim Count

- many policies have no claims (zero-inflation)
- the variance and mean of claim count is quite different (violates Poisson's assumption)
```{r echo=FALSE}
plot_ly(claims_data,
        x = ~claim_count,
        type = "histogram",
        histnorm = "percent")

mean(claims_data$claim_count)
var(claims_data$claim_count)

```


## Frequency Analysis by rating factor

### Solar System
Significantly less data is found in Hellionis Cluster, relative to the other systems. However when looking at the average claim count, the difference between the solar systems is very marginal.
```{r echo=FALSE}


# Distribution of solar system by claims
group_solar_sys <- claims_data %>%
  group_by(solar_system) %>%
  summarise(
    claims = sum(claim_count),
    count = n(),
    avg_claim_num = mean(claim_count)
  )

plot_ly(group_solar_sys) %>%
  add_bars(x = ~solar_system, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~solar_system, y = ~claims, name = "Claim Count", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Solar System vs Claim Count",
    xaxis = list(title = "Solar System"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Claim Count",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )

plot_ly(group_solar_sys) %>%
  add_bars(x = ~solar_system, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~solar_system, y = ~avg_claim_num, name = "Average Claim Count", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Solar System vs Average Claim Count",
    xaxis = list(title = "Solar System"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Average Claim Count",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )


```



### Energy Back Up
According to the data dictionary, energy backup score only takes values from 1 to 5. Therefore we will convert this variable into a factor.

```{r echo=FALSE}

group_EBS <- claims_data %>%
  group_by(energy_backup_score) %>%
  summarise(
    claims = sum(claim_count),
    count = n()
  )
plot_ly(group_EBS) %>%
  add_bars(x = ~energy_backup_score, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~energy_backup_score, y = ~claims, name = "Claim Count", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Energy Backup Score vs Claim Count",
    xaxis = list(title = "Energy Backup Score"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Claim Count",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )


```

### Supply Chain Index
```{r echo=FALSE}

plot_ly(
  claims_data,
  x = ~factor(claim_count),
  y = ~supply_chain_index,
  type = "box",
  jitter = 0.3
) %>%
  layout(
    title = "Box Plot of Supply Chain Index by Claim Count"
  )
```


### Production Load
```{r echo=FALSE}

plot_ly(
  claims_data,
  x = ~factor(claim_count),
  y = ~production_load,
  type = "box",
  jitter = 0.3
) %>%
  layout(
    title = "Box Plot of Production Load by Claim Count"
  )
```

### Average Crew Experience
```{r echo=FALSE}

group_crew_exp <- claims_data %>%
  group_by(avg_crew_exp) %>%
  summarise(
    claims = sum(claim_count),
    count = n()
  )
plot_ly(group_crew_exp) %>%
  add_bars(x = ~avg_crew_exp, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~avg_crew_exp, y = ~claims, name = "Claim Count", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Average Crew Experience vs Claim Count with Exposure",
    xaxis = list(title = "Average Crew Experience"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Claim Count",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )

```

### Safety Compliance 
```{r echo=FALSE}

group_sc <- claims_data %>%
  group_by(safety_compliance) %>%
  summarise(
    claims = sum(claim_count),
    count = n()
  )
plot_ly(group_sc) %>%
  add_bars(x = ~safety_compliance, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~safety_compliance, y = ~claims, name = "Claim Count", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Safety Compliance vs Claim Count",
    xaxis = list(title = "Safety Compliance"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Claim Count",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )


```




## Severity Analysis by Rating Factor
```{r echo=FALSE}
severity_data <- read_excel("BI Claims.xlsx", sheet = "sev")

original_row_count <- nrow(severity_data)

# Display table that shows the number of missing rows
missing_summary <- severity_data %>%
  summarise(across(everything(), ~mean(is.na(.)) * 100)) %>%  # % missing
  pivot_longer(
    cols = everything(),
    names_to = "column",
    values_to = "percent_missing"
  )

# overall number of missing rows
rows_with_na <- severity_data %>%
  filter(if_any(everything(), is.na)) %>%  # TRUE if any column in the row is NA
  nrow()

overall_percent <- rows_with_na / nrow(severity_data) * 100

# add overall row to the summary
missing_summary <- missing_summary %>%
  bind_rows(tibble(
    column = "Rows with NA",
    percent_missing = overall_percent
  ))

# Display table in R Markdown
kable(missing_summary, digits = 2, caption = "Percent of Missing Values per Column")

# cleaning solar system
severity_data <- severity_data %>%
  drop_na() %>%
  mutate(
    policy_id = sub("_.*", "", policy_id),
    station_id = sub("_.*", "", station_id),
    solar_system = sub("_.*", "", solar_system),
    claim_id = sub("_.*", "", claim_id),
    claim_seq = round(claim_seq)
  ) %>%
  filter(
    production_load >= 0,
    production_load <= 1,
    energy_backup_score %in% c(1,2,3,4,5),
    safety_compliance %in% c(1,2,3,4,5),
    exposure >= 0,
    exposure <= 1,
    claim_seq %in% c(1,2,3,4),
    claim_amount <= quantile(claim_amount, 0.995),
    claim_amount > 0
  )


# Removed Rows
paste0("Removed ", ((original_row_count - nrow(severity_data)) / original_row_count) * 100, "% of the data (defective)")
```


## Distribution of Losses
```{r echo=FALSE}
# find the 99.5% quantile
quantile.5 <- severity_data %>%
  filter(claim_amount <= quantile(claim_amount, 0.995), claim_amount > 0)

plot_ly(quantile.5,
        x = ~claim_amount,
        type = "histogram",
        histnorm = "percent")

plot_ly(quantile.5,
        x = ~log(claim_amount),
        type = "histogram",
        histnorm = "percent")

```


### Solar System
```{r echo=FALSE}


# Distribution of solar system by claims
group_solar_sys <- severity_data %>%
  group_by(solar_system) %>%
  summarise(
    claims = mean(claim_amount),
    count = n()
  )

plot_ly(group_solar_sys) %>%
  add_bars(x = ~solar_system, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~solar_system, y = ~claims, name = "Average Claim Amount", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Solar System vs Claim Amount",
    xaxis = list(title = "Solar System"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Claim Amount",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )

```


## Energy Backup Score
```{r echo=FALSE}
group_EBS <- severity_data %>%
  group_by(energy_backup_score) %>%
  summarise(
    claims = mean(claim_amount),
    count = n()
  )
plot_ly(group_EBS) %>%
  add_bars(x = ~energy_backup_score, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~energy_backup_score, y = ~claims, name = "Average Claim Amount", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Energy Backup Score vs Claim Count",
    xaxis = list(title = "Energy Backup Score"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Average Claim Amount",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )


```

### Safety Compliance
```{r echo=FALSE}

group_sc <- claims_data %>%
  group_by(safety_compliance) %>%
  summarise(
    claims = sum(claim_count),
    count = n()
  )
plot_ly(group_sc) %>%
  add_bars(x = ~safety_compliance, y = ~count, name = "Exposure (Count)", opacity = 0.3) %>%
  add_lines(x = ~safety_compliance, y = ~claims, name = "Claim Count", yaxis = "y2", line = list(color = "red")) %>%
  layout(
    title = "Safety Compliance vs Claim Count",
    xaxis = list(title = "Safety Compliance"),
    yaxis = list(title = "Exposure (Count)"),           # left y-axis for bars
    yaxis2 = list(
      title = "Claim Count",
      overlaying = "y",   # overlay on same plot
      side = "right"      # right-hand axis
    ),
    barmode = "overlay"
  )


```

### Production Load
```{r echo=FALSE}

plot_ly(
  severity_data,
  x = ~production_load,
  y = ~claim_amount,
  type = "scatter"
) %>%
  layout(
    title = "Line Graph of Production Load vs Claim Amount"
  )
```


```{r include=FALSE}
write_xlsx(severity_data, "sev_data_cleaned.xlsx")
write_xlsx(claims_data, "claim_data_cleaned.xlsx")
```

